Draft

### 2017-08-26 Thesis/Draft.Rmd
Update APT section

---
title: "Draft"
author: "Tim Radtke"
date: "5/28/2017"
header-includes:
   #- \documentclass[a4paper,twoside,english]{article}
   - \usepackage[retainorgcmds]{IEEEtrantools}
   - \usepackage{bm}
   - \usepackage{amsmath}
   - \usepackage{bbm}
   - \usepackage[lined,boxed]{algorithm2e}
   - \newtheorem{theorem}{Theorem}
   - \newtheorem{lemma}{Lemma}
   - \newcommand{\KL}{\,\text{KL}}
   - \newcommand{\der}{\,\text{d}}
   #- \usepackage[top=1.5cm, bottom=1.5cm, outer=5cm, inner=2cm, heightrounded, marginparwidth=2.5cm, marginparsep=2cm]{geometry}
   #- \usepackage{fontspec}
   #-   \setmainfont{Arial}
keep_tex: TRUE
output:
  pdf_document:
    number_sections: true
    keep_tex: true
    toc: true
    toc_depth: 2
    #latex_engine: xelatex
fontsize: 12pt
urlcolor: blue
---

@@

# Introduction to Sequential Sampling and Multi-Armed Bandits

@@

# Thresholding Bandit Problem \label{Chapter_TBM}

To introduce the thresholding bandit problem, consider a game in which the learner is tasked to find those slot machines within a row of $K$ slot machines in a casino that give a mean payoff larger than a fixed threshold $\tau$, for example $\tau = 0$. The learner cannot pull all slot machines at the same time. Instead, he acquires samples sequentially from one slot machine at a time to learn about their payoff structure. The learner has a budget of $T$ slot machine pulls. After $t-1$ samples, the learner has observed $T_i(t-1)$ samples from slot machine $i$. The payoffs $X_{i,1}, ..., X_{i,T_i(t-1)}$ are drawn from the payoff distribution $\nu_i$ underlying the slot machine $i$. At time $t$, the learner considers the payoffs he observed so far in order to choose the next slot machine he would like to pull. His decision is lead by the goal to classify the slot machines with distribution $\nu_i$ and mean payoff $\mu_i$ after $T$ samples as being above or below the threshold with error probability as small as possible.

@@

## Setup

Consider a setting of $i=1, ..., K$ unknown probability distributions $\nu_i$. To stick with the idea of multi-armed bandits, we call those distributions arms, and observations from those arms rewards (Bubeck and Cesa-Bianchi, 2012). The learner is interested in learning about the structure of the arms; in particular, he would like to estimate the means $\mu_1, ..., \mu_K$ by sequentially drawing observations $X_{i,1}, X_{i,2}, ...$ from each arm. However, at time $t$ the learner cannot observe a reward from every distribution but must select a single arm $I_t$ from which he would like to observe the reward $X_{I_t,t}$. An adaptive sampling strategy makes the decision of which arm to pull usually by computing an index $B_i(t)$ for each arm. If $T_i(t)$ is the number of samples observed until time $t$ for arm $i$, then $B_i(t)$ is a function of the past rewards $X_{i,1}, ..., X_{i,T_i(t)}$ observed for arm $i$, and w.l.o.g. $I_t = \arg \min_{i = 1,...K} B_i(t)$. The set of indices can be called strategy (Kaufmann et al., 2012) which makes it clear that the strategy is adapted as more observations are collected. Strategies of this kind are also called index policies.

What index proves to be useful is determined by the loss function with which the decisions of the algorithm are evaluated. While many problems are concerned with finding the arm with the largest mean $\mu_i$, in which algorithms collect loss either cumulatively (Auer and Cesa-Bianchi, 2012) or once at the end of the sampling (Kaufmann et al., 2016), the thresholding bandit problem considers a slightly different loss function. Consider a fixed threshold value $\tau \in \mathbb{R}$ and an epsilon interval $\tau \pm \epsilon$ around the threshold for $\epsilon \geq 0$.

------------ Old

In any case, the problem boils down to the following. As usual in multi-armed bandit problems, we have $K$ arms $\mathcal{A} = \{1, \dots K\}$. We can pull arm $k$ at time $t$ to collect feedback in form of the random variable $X_{k,t}$ which is distributed according to the arm's distribution $\nu_k$. In general, we are concerned with estimating for each arm $i$ the mean $\mu_i$ of its distribution $\nu_i$. In contrast to most bandit problems, however, we do not directly compare arms with each other by comparing their means. That would for example be the case for pure exploration bandits, where it is necessary to compare the arms because one aims to find the best one relative to the others. In the case of the thresholding bandit, however, we compare each arm's mean $\mu_i$ individually against the threshold $\tau \in \mathbb{R}$ which is known and fixed before the experiment. One might compare this to a pure exploration bandit in which the mean of the best arm is known upfront. We will see later that this fact leads to advantages in the design of algorithms which are not as readily available in other bandit settings.

In Locatelli (2016), the arms' distributions are assumed to follow an $R$-sub-Gaussian distribution. Under this assumption, the authors are able to proof a general lower bound as well as an algorithm with a matching upper bound. Both hinge upon the *complexity* of a given problem. Since bandit problems can differ in the number of the arms $K$ and the means of each arm, some pose easier identification problems than others. Indeed, in the thresholding problem, the difficulty of the problem naturally depends on how close (the means of the) distributions are to the fixed threshold. Intuitively, the closer a mean $\mu_i$ is to the threshold $\tau$, the more difficult it is to distinguish the two successfully based on samples from the distribution $\nu_i$. A straightforward measure of this distance is given by $\Delta_i = |\tau - \mu_i| + \epsilon$, where $\epsilon$ defines a $2\epsilon$ interval around the threshold $\tau$ in which we allow ourselves to make a mistake at no cost. In the case of sub-Gaussian distributions, this distance is not only simple but can be interpreted very naturally as we will come to understand later.

If we extend this idea to the set of arms $i \in \{1,...,K\}$ in a given bandit problem, a measure of the overall diffulty of the problem, or the problem's complexity, might be given by $H = \sum_{i=1}^{K} \Delta_i^{-2}$. In particular, it shows that just adding additional arms leads to an increase in complexity (since we need a larger budget to sample each arm the same number of times as before). Likewise, a smaller $\Delta_i$ increases the complexity as arms become harder to classify. Choosing the complexity measure in this way fits the distance measure $\Delta_i$ and turn out to be useful in the proof of an upper bound for the APT algorithm presented in Locatelli et al. (2016).

We also see that this definition may lose relevancy as we move away from $R$-sub-Gaussian distributions. (ADD QUICK EXAMPLE OF WHAT CHANGES; REFER TO LATER PART AND DESCRIBE THIS AS THE MOTIVATION OF THE PAPER)

Another characteristic uniquely identifying bandit problems is the regret measure that we aim to optimize. As explained during the introduction, multi-armed bandits traditionally optimize cumulative regret. The notion of regret considered in the thresholding bandit, however, is more akin to the simple regret in pure exploration bandits or the error probability in sequential hypothesis testing. Indeed, one can most easily describe the regret in the thresholding case as the probability of misclassifying any arm. Or, more rigorously, define the set of all arms with mean larger than the threshold, $\mathcal{S}_\tau = \{i \in \mathcal{A}: \mu_i \geq \tau\}$. Consequently, $\mathcal{S}_\tau^C = \{i \in \mathcal{A}: \mu_i < \tau\}$. At the end of the sampling rule/procedure (!!we have not mentioned yet the fact that we're in the fixed budget case!!), we demand that any algorithm returns the two corresponding *estimated* sets: $\hat{\mathcal{S}}_{\tau} = \{i \in \mathcal{A}: \hat{\mu}_i \geq \tau\}$, as well as $\hat{\mathcal{S}}_{\tau}^C = \{i \in \mathcal{A}: \hat{\mu}_i < \tau\}$. We simply demand an answer to the question which arms have means above the threshold $/tau$, and which arms lie below  $\tau$. Let $T$ be the overall number of samples the algorithm is allowed to pull. We define the loss as $\mathcal{L}(T) = \mathbbm{1}_{\{\hat{\mathcal{S}}_{\tau} \cap \mathcal{S}_{\tau-\epsilon}^C \neq \emptyset\}} \cdot \mathbbm{1}_{\{\hat{\mathcal{S}}_{\tau}^C \cap \mathcal{S}_{\tau+\epsilon} \neq \emptyset\}}$. That is, we compare the return set with the true set and collect a loss of 1 if any arm with a true mean outside the $2\epsilon$ interval around $\tau$ was misclassified. It is important to realize that arms with means inside the $2 \epsilon$ interval may be misclassified without it affecting the loss. These arms are considered close enough to the threshold that we do not care whether they lie above or below.

ADD A GRAPHIC THAT VISUALIZES THE ESTIMATED AND THE REQUIRED RETURN SETS

Now, the expected regret/loss is defined as $\mathbb{E}[\mathcal{L}(T)] = \mathbb{P}((\hat{\mathcal{S}}_{\tau} \cap \mathcal{S}_{\tau-\epsilon}^C \neq \emptyset) \land (\hat{\mathcal{S}}_{\tau}^C \cap \mathcal{S}_{\tau+\epsilon} \neq \emptyset))$. This, of course, is the mentioned probability of making a mistake.

Lastly, it is to mention that we are considering the fixed-budget version of the thresholding bandit problem. We thus fix the number of samples that the algorithm will pull upfront at $T$. The task of the algorithm is then to find a sampling rule or strategy which distributes the budget $T$ among the $K$ arms in such a way that the expected regret, or the error probability, is minimized. In other terms, the algorithm aims to use the $T$ samples in a way that lets him return the two sets $\hat{\mathcal{S}}_{\tau}$ and $\hat{\mathcal{S}}_{\tau}^C$ with as much confidence as possible. This again shows the contrast to the fixed confidence setting in which one fixes the confidence level upfront and lets the algorithm draw as many samples as necessary to reach the required confidence.

At this point, we have defined the fixed budget thresholding bandit setup for any algorithm in the case of sub-Gaussian distributions. We thus proceed to describe the APT algorithm proposed in Locatelli et al. (2016).

## The APT Algorithm

\IncMargin{1em}
\begin{algorithm}
\SetKwData{Left}{left}\SetKwData{This}{this}\SetKwData{Up}{up}
\SetKwFunction{Union}{Union}\SetKwFunction{FindCompress}{FindCompress}
\SetKwInOut{Input}{Input}\SetKwInOut{Output}{Output}
\Input{$\tau$, $\epsilon$}
\Output{$\hat{S}_{\tau} = \{k: \hat{\mu}_k(T) \geq \tau\}$}
\BlankLine
\emph{Pull each arm once}\;
\For{$t = K+1$ \KwTo $T$}{
  \emph{Pull arm $I_t = \arg \min_{k\leq K} B_k^{APT}(t) = \sqrt{T_k(t)} \sqrt{|\hat{\mu}_i(t) - \tau|+\epsilon}$}\;
  \emph{Observe reward $X \sim \nu_{I_t}$}\;
}
\caption{Anytime Parameter-free Thresholding Algorithm (Locatelli et al., 2016)}\label{algo_apt}
\end{algorithm}\DecMargin{1em}

In the following we describe the *Anytime Parameter-free Thresholding* (APT) algorithm proposed in Locatelli et al. (2016), which is a simple yet effective algorithm and displays some very favorable characteristics when compared with alternative algorithms.

From the general setup, we already know how the algorithm will end. After it has drawn $T$ samples, the algorithm makes a decision of which arms to include in the return sets on basis of the empirical means $\hat{\mu}_i(T)$ estimated based on the samples drawn until time $T$. Furthermore, the threshold and interval around the threshold are fixed upfront and serve as argument to the algorithm: $(\tau, \epsilon)$. For a given arm $i$ and a given time $t$, we let $T_i(t)$ represent the number of samples drawn from arm $i$ until time $t$ (including). Consequently, the empirical mean of arm $i$ at time $t$ is given by $\hat{\mu}_i(t) = \frac{1}{T_i(t)} \sum_{s=1}^{T_i(t)} X_{i,s}$ where $X_{i,s}$ is the feedback sampled from distribution $\nu_i$ when the arm is pulled for the $s$-th time.

To be specific, the APT algorithm initiates by pulling each arm once. Then, for each $t > K$, it calculates for every arm the values $T_i(t)$ and $\mu_i(t)$. Based on the empirical mean, the algorithm computes the estimated gap between the mean and the threshold as $\hat{\Delta}_i(t) = \hat{\Delta}_i^{\tau, \epsilon}(t) = |\mu_i(t) - \tau| + \epsilon$. The decision of which arm is sampled next is then based on the value $B_i(t+1) = \sqrt{T_i(t)} \hat{\Delta}_i(t)$: the algorithm pulls arm $I_{t+1} = \arg \min_{i\leq K} B_i(t+1)$ minimizing this decision quantity. After $T$ pulls overall, the arm returns the sets described in the general setup, $\hat{\mathcal{S}}_{\tau}$ and $\hat{\mathcal{S}}^C_{\tau}$.

The APT algorithm tends to pull arms more often with corresponding mean close to the threshold and interval (small $\hat{\Delta}_i(t))$. This makes sure that more evidence is gathered on arms which appear to be inherently difficult to classify based on the samples collected until time $T_i(t)$, as these arms stifle the confidence the most. That is, until we have collected enough evidence to make a confident classification even if the mean is truly close to the threshold. In that case, $\sqrt{T_i(t)}$ will raise $B_i(t+1)$. Consequently, other arms farther away from the threshold will be sampled again. In this way, the algorithm tries to distribute the budget $T$ over all arms so as to define a sampling rule optimizing the confidence of the final decision.

Locatelli et al. (2016) show that the APT algorithm is optimal in the sense that the upper bound on its expected loss reaches the lower bound for the expected loss.

>>>>

---
title: "Draft"
header-includes:
   #- \documentclass[a4paper,twoside,english]{article}
   - \usepackage[retainorgcmds]{IEEEtrantools}
   - \usepackage{bm}
   - \usepackage{amsmath}
   - \usepackage{bbm}
   - \usepackage{hyperref}
   - \usepackage[lined,boxed]{algorithm2e}
   - \newtheorem{theorem}{Theorem}
   - \newtheorem{definition}{Definition}
   - \newtheorem{lemma}{Lemma}
   - \newcommand{\KL}{\,\text{KL}}
   - \newcommand{\der}{\,\text{d}}
   #- \usepackage[top=1.5cm, bottom=1.5cm, outer=5cm, inner=2cm, heightrounded, marginparwidth=2.5cm, marginparsep=2cm]{geometry}
   #- \usepackage{fontspec}
   #-   \setmainfont{Arial}
keep_tex: TRUE
output:
  pdf_document:
    number_sections: true
    keep_tex: true
    toc: true
    toc_depth: 2
    #latex_engine: xelatex
fontsize: 12pt
urlcolor: blue
---

@@

# Introduction \label{chap:Introduction}

@@

# Thresholding Bandit Problem \label{chap:ThresholdingBanditProblem}

To introduce the thresholding bandit problem, consider a game in which the learner is tasked to find those slot machines within a row of $K$ slot machines in a casino that give a mean payoff larger than a fixed threshold $\tau$, for example $\tau = 0$. The learner cannot pull all slot machines at the same time. Instead, he acquires samples sequentially from one slot machine at a time to learn about their payoff structure. The learner has a budget of $T$ slot machine pulls. After $t-1$ samples, the learner has observed $T_i(t-1)$ samples from slot machine $i$. The payoffs $X_{i,1}, ..., X_{i,T_i(t-1)}$ are drawn from the payoff distribution $\nu_i$ underlying the slot machine $i$. At time $t$, the learner considers the payoffs he observed so far in order to choose the next slot machine he would like to pull. His decision is lead by the goal to classify the slot machines with distribution $\nu_i$ and mean payoff $\mu_i = \mathbb{E}_{X \sim \nu_i}[X]$ after $T$ samples as being above or below the threshold with error probability as small as possible.

@@

## Setup \label{sec:Setup}

Consider a setting with a set of $\mathbb{A}=\{1, ..., K\}$ unknown probability distributions $\nu_i, i \in \mathbb{A}$. To stick with the idea of multi-armed bandits, we call those distributions arms, and observations from those arms rewards (Bubeck and Cesa-Bianchi, 2012). The learner is interested in learning about the structure of the arms; in particular, he would like to estimate the means $\mu_1, ..., \mu_K$ by sequentially drawing observations $X_{i,1}, X_{i,2}, ...$ from each arm. However, at time $t$ the learner cannot observe a reward from every distribution but must select a single arm $I_t$ from which he would like to observe the reward $X_{I_t,t}$. An adaptive sampling strategy makes the decision of which arm to pull usually by computing an index $B_i(t)$ for each arm. If $T_i(t)$ is the number of samples observed until time $t$ for arm $i$, then $B_i(t)$ is a function of the past rewards $X_{i,1}, ..., X_{i,T_i(t)}$ observed for arm $i$, and w.l.o.g. $I_t = \arg \min_{i = 1,...K} B_i(t)$. The set of indices can be called strategy (Kaufmann et al., 2012) which makes it clear that the strategy is adapted as more observations are collected. Strategies of this kind are also called index policies.

What index proves to be useful is determined by the loss function with which the decisions of the algorithm are evaluated. While many problems are concerned with finding the arm with the largest mean $\mu_i$, in which algorithms collect loss either cumulatively (Auer and Cesa-Bianchi, 2012) or once at the end of the sampling (Kaufmann et al., 2016), the thresholding bandit problem considers a slightly different loss function. Consider a fixed threshold value $\tau \in \mathbb{R}$ and an epsilon interval $\tau \pm \epsilon$ around the threshold for $\epsilon \geq 0$. As mentioned in the introduction, the learner is tasked with classifying arms as having a mean $\mu_i$ larger or smaller than $\tau$. To describe which arms are above the threshold, we introduce the following set notation (compare Locatelli et al., 2016): Let $\mathcal{S}_{\tau}$ be the set of distributions in $mathbb{A}$ with mean larger than the threshold: $\mathcal{S}_{\tau} := \{i \in \mathbb{A}, \mu_i \geq \tau\}$. Consequently, the complement is given by the arms with mean below the threshold, $\mathcal{S}^C_{\tau} := \{i \in \mathbb{A}, \mu_i < \tau\}$.

The thresholding problem for the learner is now to use a budget of overall $T$ samples (or rounds) to explore the different arms. After this exploration phase, the learner has to return a set of arms that he believes to have a mean above the threshold. The problem can thus be described by the threshold $\tau$ and a precision $\epsilon$. When evaluating the classification of the learner, arms with mean inside of the epsilon interval will not be considered. The learner only has to distinguish the means from the threshold correctly up to the precision $\epsilon$. Thus, the algorithm is not evaluated against $\mathcal{S}_{\tau}$ and $\mathcal{S}^C_{\tau}$, but needs to distinguish $\mathcal{S}_{\tau+\epsilon}$ from $\mathcal{S}^C_{\tau-\epsilon}$. As in Locatelli et al. (2016), whenever we say that "an arm is over the threshold", we mean that the mean of the arm is over the threshold.

The learner's classification is returned as the set $\hat{\mathcal{S}}_{\tau} := \hat{\mathcal{S}}_{\tau}(T) \subset \mathbb{A}$ after $T$ rounds; that is, the set of arms estimated to be over the threshold. The classification is evaluated against the loss given as (Locatelli et al., 2016):

$$
\mathcal{L}(T) = \mathbbm{1}{(\hat{\mathcal{S}}_{\tau} \cap \mathcal{S}_{\tau-\epsilon}^C \neq \emptyset) \lor (\hat{\mathcal{S}}_{\tau}^C \cap \mathcal{S}_{\tau+\epsilon} \neq \emptyset)}
$$

In order to perform well, a learner has to identify all arms that are over the threshold as such, and to not claim arms as being over the threshold if they actually are below the threshold. Due to the precision $\epsilon$, the loss is not influenced by arms which are too close to the threshold $\tau$, that is, arms inside of the $\tau \pm \epsilon$ interval. If either $\hat{\mathcal{S}}_{\tau}$ contains an arm below $(\tau - \epsilon)$, or $\hat{\mathcal{S}}^C_{\tau}$ contains an arm above $(\tau + \epsilon)$, a loss of $1$ is accrued. Else, all arms outside the epsilon interval are classified correctly and the loss is equal to $0$. In the case of $\epsilon = 0$, all arms need to be classified correctly to not endure any loss.

Algorithms for this thresholding problem can be compared based on their expected loss. Let $\mathbb{E}$ be the expectation with regard to the samples $X_{i,1}, ..., X_{i,T}$ collected by the algorithm. The expected loss is then given as

$$
\mathbb{E}[\mathcal{L}(T)] = \mathbb{P}\big((\hat{\mathcal{S}}_{\tau} \cap \mathcal{S}_{\tau-\epsilon}^C \neq \emptyset) \lor(\hat{\mathcal{S}}_{\tau}^C \cap \mathcal{S}_{\tau+\epsilon} \neq \emptyset)\big),
$$

which corresponds to the probability of making a wrong classification. Both the probability of not identifying an arm that is actually over the threshold, as well as the probability of accepting an arm as over the threshold when it is not are evaluated. The oracle strategy would evaluate all arms correctly with an error probability of $0$ (Locatelli et al., 2016).

At this point, we have introduced the general thresholding bandit problem setting which we will use throughout the thesis. The setting might be extended by different assumptions to make certain algorithms or bounds feasible. For that reason, Locatelli et al. (2016) assume that the distributions $\nu_i, i \in \mathbb{A}$ are $R$-sub-Gaussian. Under this assumption, the authors are able to prove a lower bound on the largest probability of error of the best algorithm in the setting we have just introduced, as well as an upper bound on the performance of their APT algorithm we will introduce in the following chapter. As in their paper, we define $R$-sub-Gaussian and assume it in the description of their algorithm and bounds:

\begin{definition}[$R$-sub-Gaussian distribution] 
\label{definition:rsubgaussian}
Let $R > 0$. A distribution $\nu$ is $R$-sub-Gaussian if for all $t \in \mathbb{R}$ we have
\begin{equation*}
\mathbb{E}_{X \sim \nu}[\exp(tX - t\mathbb{E}[X])] \leq \exp(R^2t^2/2).
\end{equation*}
\end{definition}

As noted by Locatelli et al. (2016), this ensures that the mean $\mu_i$ of the distributions $\nu_i$ considered in our problem are finite. $R$-sub-Gaussian distributions furthermore offer concentration inequalities for the concentration of the empirical mean to the true mean, which makes it possible to deduce the bounds presented by the authors and repeated below. The family of $R$-sub-Gaussian distributions encompasses Gaussian distributions of variance $R^2$ for $R \in \mathbb{R}$, as well as all bounded distributions, with the important case of Bernoulli distributions.

Having described the thresholding bandit setting as well as the assumption of $R$-sub-Gaussianity, we can now proceed to describe the *Anytime Parameter-free Thresholding* (APT) algorithm proposed in Locatelli et al. (2016) for the thresholding bandit problem.

## The Anytime Parameter-free Thresholding Algorithm \label{sec:TheAPTAlgorithm}

In the following we describe the *Anytime Parameter-free Thresholding* (APT) algorithm proposed in Locatelli et al. (2016), which is a simple yet effective algorithm and displays some very favorable characteristics when compared to alternative algorithms.

We fix the threshold $\tau$ and epsilon interval around the threshold given by $\epsilon$ upfront, and supply them as arguments to the algorithm: $(\tau, \epsilon)$. Based on the general thresholding problem, we demand that the algorithm draws samples $X_{i,t}$ for $T$ rounds from the distributions $\nu_i, i\in \mathbb{A}$. For a given arm $i$ and a given time $t$, we let $T_i(t)$ represent the number of samples drawn from arm $i$ until time $t$ (including). Consequently, the empirical mean of arm $i$ at time $t$ is given by $\hat{\mu}_i(t) = \frac{1}{T_i(t)} \sum_{s=1}^{T_i(t)} X_{i,s}$ where $X_{i,s}$ is the feedback sampled from distribution $\nu_i$ when the arm is pulled for the $s$-th time. In the case of the APT algorithm, the decision $\hat{\mathcal{S}}_{\tau}$ is based on the empirical means $\hat{\mu}_i(T)$ estimated from the samples drawn up to time $T$: $\hat{\mathcal{S}}_{\tau} := \hat{\mathcal{S}}_{\tau}(T) = \{i \in \mathbb{A}, \hat{\mu}_i(T) \geq \tau\}$.

As we see in algorithm \ref{alg:APT}, the APT algorithm initiates by pulling each arm once. Then, for each $t > K$, it calculates for every arm the values $T_i(t)$ and $\hat{\mu}_i(t)$. Based on the empirical mean, the algorithm computes the estimated gap between the mean and the threshold as $\hat{\Delta}_i(t) = \hat{\Delta}_i^{\tau, \epsilon}(t) = |\mu_i(t) - \tau| + \epsilon$. The decision of which arm is sampled next is then based on the value $B_i(t+1) = \sqrt{T_i(t)} \hat{\Delta}_i(t)$: the algorithm pulls arm $I_{t+1} = \arg \min_{i\leq K} B_i(t+1)$ minimizing this decision quantity. After overall $T$ pulls, the algorithm returns the sets described in the general setup, $\hat{\mathcal{S}}_{\tau}$ and $\hat{\mathcal{S}}^C_{\tau}$.

\IncMargin{1em}
\begin{algorithm}
\SetKwData{Left}{left}\SetKwData{This}{this}\SetKwData{Up}{up}
\SetKwFunction{Union}{Union}\SetKwFunction{FindCompress}{FindCompress}
\SetKwInOut{Input}{Input}\SetKwInOut{Output}{Output}
\Input{$\tau$, $\epsilon$}
\Output{$\hat{S}_{\tau} = \{k: \hat{\mu}_k(T) \geq \tau\}$}
\BlankLine
\emph{Pull each arm once}\;
\For{$t = K+1$ \KwTo $T$}{
  \emph{Pull arm $I_t = \arg \min_{k\leq K} B_k^{APT}(t) = \sqrt{T_k(t)} (|\hat{\mu}_i(t) - \tau|+\epsilon)$}\;
  \emph{Observe reward $X \sim \nu_{I_t}$}\;
}
\caption{Anytime Parameter-free Thresholding Algorithm (Locatelli et al., 2016)}\label{alg:APT}
\end{algorithm}\DecMargin{1em}

Just like the general index policies mentioned in \autoref{sec:Setup}, we see that the APT algorithm adapts its strategy using an index $B_i(t+1)$ for each arm that is based entirely on the observations observed until time $t$. Given the arm with the smallest index is pulled at round $t+1$, we observe that arms that are estimated to be further away from the threshold $\tau$ as measured by the absolute distance to their empirical mean $\hat{\mu}_i$, $\hat{\Delta}_i(t)$, are pulled less often. Intuitively, it is easier to classify arms as below or above $\tau$ when their mean is far away from the threshold. In contrast to the uniform sampling strategy, APT will gather more evidence for arms which appear to be inherently more difficult to classify. Likewise, APT tends to pull arms less often which it already pulled considerably often as measured by $T_i(t)$. If arm $i$ has been pulled more often than arm $j$, then we should in general be more confident in its estimated mean and the corresponding distance to the threshold. If we are confident in $i$, then it is better to explore other arms by drawing an observation from arm $j$. By adapting its sampling strategy in this way during the sequential sampling process, the APT algorithm minimizes the probability of error $\mathbb{E}[\mathcal{L}(T)]$ after $T$ samples have been observed.

The previous paragraph describes how arms close to the threshold are intuitively more difficult to classify. In the index employed by the APT algorithm, the distance between an arm and the threshold is measured by the absolute distance $\hat{\Delta}_i(t)$ between the (estimated) mean of the arm and the threshold $\tau$ (and the epsilon interval). Given that one compares $B_i^{APT}$ with $B_j^{APT}$, one could just as well measure the distance by the squared distance, $\hat{\Delta}_i(t)^2$. When the APT algorithm realizes that an arm has a mean close to the threshold, then this arm requires more attention than an arm far away from the threshold; the confidence in the empirical mean has to be larger. Or the empirical mean has to be better concentrated around the true mean. Otherwise it is easier to make a mistake on this arm. To gain this confidence, the arm requires more samples, and thus a larger share of the overall $T$ samples available. With this intuition, it makes sense to measure the complexity of a thresholding problem: If we knew upfront how complex a given set of arms is, we knew whether we can expect a small or large probability of error given our budget $T$. Also, for example in the best arm identification problem, there exist algorithms which take the complexity of a problem as input. Based on the complexity they then adapt their sampling strategy. See for example the UCB-E algorithm in Audibert et al. (2010) for the best arm identification problem, or the KL-LUCB-E algorithm in Kaufmann and Kalyanakrishnan (2013) as extension to the Top-$m$ problem.

While the APT algorithm does not require the complexity to be known upfront (which would be problematic), it is necessary to define the complexity to deduce the upper bound on its probability of error. Similar to the distance measure used in the algorithm's index, we define the gap $\Delta_i^{\tau,\epsilon}$ between arm $i$ and the threshold $\tau \pm \epsilon$ as

$$
\Delta_i := \Delta_i^{\tau,\epsilon} = |\mu_i - \tau| + \epsilon
$$

Given this gap, Locatelli et al. (2016) define the complexity $H_{\tau, \epsilon}$ for the thresholding bandit problem naturally as

$$
H := H_{\tau, \epsilon} = \sum_{i=1}^K \frac{1}{(\Delta_i^{\tau,\epsilon})^2}.
$$

Note that the complexity of the problem increases in the number of arms $K$ and decreases in the gap $\Delta_i^{\tau,\epsilon}$. Consequently, when we require a smaller precision, that is, we increase $\epsilon$, the hardness decreases. Furthermore, we see that each arm contributes its individual complexity, $h_i^{\tau,\epsilon} = \frac{1}{(\Delta_i^{\tau,\epsilon})^2}$ linearly into the problem. Thus, if the complexity $H$ is a good measure of the actual problem complexity, and we knew upfront which arm has which mean, we could allocate the "optimal" share of the budget, $T \cdot h_i /H$ to arm $i$. However, since we do not know neither $\mu_i$ nor $H$, the best we can do is to adapt our allocation of samples more and more towards this optimum as we observe samples.

At this point, one can question whether the measure of complexity $H$ proposed in Locatelli et al. (2016) is a good measure of the "true" problem complexity. In the following \autoref{sec:UpperBoundLocatelli}, we will show the upper bound on the probability of error for the APT algorithm and see how the complexity $H$ arises naturally for the APT algorithm. In \autoref{sec:LowerBoundLocatelli} we then show the lower bound presented in Locatelli et al. (2016) and that the APT algorithm is optimal in that it matches the lower bound. We discuss the results and the APT algorithm further in \autoref{sec:DiscussionLocatelli}.

## Upper Bound for the APT Algorithm \label{sec:UpperBoundLocatelli}

In this chapter we present the upper bound on the APT algorithm proven in Locatelli et al. (2016). The upper bound is on the probability of error that the algorithm has after $T$ rounds of sampling for a problem with arms that are assumed to be $R$-sub-Gaussian. The bound depends also on the complexity of the problem, and we show in the proof that given the index of the APT algorithm as defined in \autoref{sec:TheAPTAlgorithm}, the complexity $H$ is an appropriate measure.

We now state the theorem as in Locatelli et al. (2016) and present a proof.

\begin{theorem}[Theorem 2, Locatelli et al., 2016] \label{theorem:LocatelliTheorem4}
Let $K \geq 0$, $T \geq 2K$, and consider a problem $\mathcal{B}$. Assume that all arms $\nu_i$ of the problem are R-sub-Gaussian with means $\mu_i$. Let $\tau \in \mathbb{R}, \epsilon \geq 0$. Algorithm APT's expected loss is upper bounded on this problem as 
\begin{equation*} \mathbb{E}(\mathcal{L}(T)) \leq \exp
(-\frac{1}{64R^2}\frac{T}{H} + 2 \log (T) + 1)K)) \end{equation*}

### Proof of Theorem \ref{theorem:LocatelliTheorem4}

In order to upper bound the expected loss of the APT algorithm, we want to define an event on which the algorithm correctly seperates arms from the threshold, that is, a separation condition holds on this event. We then want to show that this *favorable* event does indeed hold with large probability. Consequently, APT's probability of error will correspond to the probability that the favorable event does not hold, as this leads to a failure of the seperation condition, and consequently to an incorrect classification of at least one arm by the algorithm. We start by defining the separation condition.

As noted in Zhong et al. (2017), a simple requirement to correctly classify arms at the final round is the following. For every arm $i \in \{1, ..., K\}$ after the final round $T$, require
$$
\hat{\mu}_{i,T_i(n)} \geq \mu_i - \frac{\mu_i - \tau + \epsilon}{2} \quad \text{if} \quad \mu_i \geq \tau
$$
or
$$
\hat{\mu}_{i,T_i(n)} \leq \mu_i + \frac{\tau - \mu_i + \epsilon}{2} \quad \text{if} \quad \mu_i < \tau
$$
Or simply in terms of the gap between the arm and the threshold as used in the index by the APT algorithm:
$$
| \hat{\mu}_{i, T_i(n)} - \mu_i | \leq \frac{|\mu_i - \tau| + \epsilon}{2} = \frac{\Delta_i}{2}.
$$

Intuitively, this condition makes sense as it ensures that the current mean estimate is closer to the true mean than to the threshold. Consequently, there is no arrangement possible in which the arm is classified wrong while the separation condition holds. Assuming $\mu_i < \tau$ and $\epsilon = 0$, an incorrect classification would require $\mu_i < \tau \leq \hat{\mu_i}$. But then  $| \hat{\mu}_{i, T_i(n)} - \mu_i | \geq |\mu_i - \tau|$, which violates the separation condition.

Thus, we need to show that this condition holds for every arm $i \in \{1,...,K\}$ with large probability after the final round $T$. In different words, we need to derive an upper bound on the probability with which this condition does not hold.

The separation condition describes what event needs to hold in order for the APT algorithm to make a correct classification: The smaller the deviation of the empirical mean from the population mean, the more difficult it is for the classification to be wrong. Moreover, if the mean and empirical mean are identical, then the classification is correct independent of the threshold. Also, the closer the mean to the threshold, the better the estimate has to be for a correct classification to be likely. To bound the absolute deviation of the empirical mean from the true mean, we define the favorable event $\xi$ as

The probability that this favorable event occurs should be large, or equivalently, the probability that the deviation is small should be large. Then it follows that the separation condition holds with large probability.

- Show Chernoff/Hoeffding bound detailed
- Show union bound to get deviation for all arms $i \leq K$ and every round $s \leq T$

We now have that with large probability the favorable event holds. On the favorable event, we have for every arm $i$:

The deviation of arm $i$ at time $t$ depends on how large $H$, $T$, and $T_i(t)$ are. While $t$, $H$, $T$ are deterministic, the number of times arm $i$ has been pulled at round $t$, $T_i(t)$, is a random variable. In order to know how large the probability of the favorable event is (and thus the error probability), we need to know the value of $T_i(t)$, or at least bound a range of potential values. Put differently: To ensure a deviation smaller than $\sqrt{\frac{T\delta^2}{HT_i(t)}}$ for arm $i$, we need to ensure that arm $i$ has been pulled at least $T_i(t)$ times. As usual in multi-armed bandits, and different from uniform sampling, the number of pulls is a random variable that also depends on how often other arms are pulled. In general, we do not want each arm to be pulled an equal amount, but each arm should be pulled some times, and not too many times.

Given that we do not follow a uniform sampling strategy, we know that at time $T$, some arms will have been pulled more frequently than others. In particular, some arms will have been pulled more than $T/K$ times. How often an arm is pulled can be related to the individual arm's contribution to the overall complexity measure. Let $h_i$ be a measure of arm $i$'s individual problem complexity and write for the overall problem complexity $H = \sum_{i = 1}^{K} h_i$. Then arm $i$ contributes a share of $h_i/H$ to the overall problem complexity. We can now show that there exists an arm that has been pulled at least as often as the share of his complexity contribution.

Given the initialization of the APT algorithm, every arm is pulled once initially. Consequently, there is a budget of $T-K$ rounds left after the initialization. As shown in Locatelli et al. (2016), there exists an arm $k$ that has been pulled after the initialization ($\geq 2$ pulls) and for which at the final round $T$ it holds: $T_k(T) - 1 \geq \frac{(T-K)h_k}{H}$. This can be shown by a contradiction. Assume that the opposite holds, that is
